\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-14 tex/paper.tex
modified section title and fix typos

Algorithm~\ref{alg:avo} represents the simplest version of AVO; however, the
variance of the noisy estimator of the gradients may be too large to be useful in many problems. Improvements are possible using variance reduction techniques such as Rao-Blackwellization,
Control Variates, and quasi-Monte Carlo~\citep{ranganath2014black, tran2017variational}. 
Similarly, various fast algorithms for calculating the exact (entropically regularized) Wasserstein distance on
empirical distributions are being developed~\citep{cuturi2013sinkhorn, genevay2016stochastic, montavon2016wasserstein}. 
Similarly, the algorithm could be adopted to use estimates to the KL distance~\citep{perez2009estimation} without the adversarial optimization for the critic. These variants may be particularly useful for small $N$.

\subsection{Empirical Bayes through Adversarial Variational Inference}

The variational objectives \ref{eqn:vo-ud}-\ref{eqn:vo-ug}
have the effect of replacing the modeled data distribution of Eqn.~\ref{eqn:p_theta} with
the parameterized marginal distribution of the generated data
\begin{equation}
    \qxpsi = \int  p(\mathbf{x}|\bftheta) q(\bftheta|\bfpsi) d\bftheta.
\end{equation}
We can think of $q(\bfx|\bfpsi)$ as a \textit{variational program} as described
in \citep{2016arXiv161009033R}, though more complicated than a simple
reparameterization of normally distributed noise $\bfz$ through a differentiable
function. In our case, the variational program is a
marginalized, non-differentiable  simulator.  Its density is  intractable;
nevertheless, it can generate samples for $\bfx$ and is differentiable with
respect to $\bfpsi$.
Operationally, we sample from this marginal model via
\begin{equation}\label{eqn:p_psi}
    \mathbf{x} \sim \qxpsi \equiv \bftheta \sim q(\bftheta|\bfpsi), \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta).
\end{equation}

@@

We can also view the optimization  through the lens of empirical Bayes (EB),
where one specifies a family of priors and $p(\bftheta | \bfpsi)$ and then estimates $\bfpsi^*$ using the data.
%In our case, we find the $\bfpsi^*$ that minimizes the Wasserstein distance between the data distribution and the 
%posterior predictive distribution. 
Traditionally, we would have access to the likelihood and optimize with respect to the family of posteriors
 $p(\bftheta | \bfx, \bfpsi) = p(\bfx | \bftheta) p(\bftheta | \bfpsi^*) / \int p(\bfx | \bftheta) p(\bftheta | \bfpsi^*) d\bftheta $
as an intermediate step to computing $\bfpsi^*$. In our case, the likelihood is intractable and we have not explicitly 
specified a prior family. However, the procedure corresponds to empirical Bayes if we identify 
$q(\bftheta | \bfpsi^*) = p(\bftheta | \bfx , \bfpsi)$. Thus we call our method AVO-EB and note that it is applicable even outside of the likelihood-free context.

Algorithm~\ref{alg:avo}, can be easily modified to have different mini-batch $M_r$ for the observed data and $M_g$ for the simulator. 
When $M_g \gg M_r$, then the variance of $\bftheta^*$ goes like $M_r$~\citep{bernton2017inference}. A final $\sqrt{N/M}$ scaling to the log-posterior $q(\bftheta | \bfpsi)$ is needed to reflect the full strength of the data.

>>>>

Algorithm~\ref{alg:avo} represents the simplest version of AVO; however, the
variance of the noisy estimator of the gradients may be too large to be useful in many problems. Improvements are possible using variance reduction techniques such as Rao-Blackwellization,
Control Variates, and quasi-Monte Carlo~\citep{ranganath2014black, tran2017variational}. 
Similarly, various fast algorithms for calculating the exact (entropically regularized) Wasserstein distance on
empirical distributions are being developed~\citep{cuturi2013sinkhorn, genevay2016stochastic, montavon2016wasserstein}. 
Moreover, the algorithm could be adopted to use estimates to the KL distance~\citep{perez2009estimation} without the adversarial optimization for the critic. These variants may be particularly useful for small $N$.

\subsection{Empirical Bayes through  Variational Inference}

The variational objectives \ref{eqn:vo-ud}-\ref{eqn:vo-ug}
have the effect of replacing the modeled data distribution of Eqn.~\ref{eqn:p_theta} with
the parameterized marginal distribution of the generated data
\begin{equation}
    \qxpsi = \int  p(\mathbf{x}|\bftheta) q(\bftheta|\bfpsi) d\bftheta.
\end{equation}
We can think of $q(\bfx|\bfpsi)$ as a \textit{variational program} as described
in \citep{2016arXiv161009033R}, though more complicated than a simple
reparameterization of normally distributed noise $\bfz$ through a differentiable
function. In our case, the variational program is a
marginalized, non-differentiable  simulator.  Its density is  intractable;
nevertheless, it can generate samples for $\bfx$ whose expectations are differentiable with
respect to $\bfpsi$.
Operationally, we sample from this marginal model via
\begin{equation}\label{eqn:p_psi}
    \mathbf{x} \sim \qxpsi \equiv \bftheta \sim q(\bftheta|\bfpsi), \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta).
\end{equation}

@@

We can also view the optimization  through the lens of empirical Bayes (EB),
where one specifies a family of priors and $p(\bftheta | \bfpsi)$ and then estimates $\bfpsi^*$ using the data.
%In our case, we find the $\bfpsi^*$ that minimizes the Wasserstein distance between the data distribution and the 
%posterior predictive distribution. 
Traditionally, we would have access to the likelihood and optimize with respect to the family of posteriors
 $p(\bftheta | \bfx, \bfpsi) = p(\bfx | \bftheta) p(\bftheta | \bfpsi) / \int p(\bfx | \bftheta) p(\bftheta | \bfpsi) d\bftheta $
as an intermediate step to computing $\bfpsi^*$. In our case, the likelihood is intractable and we have not explicitly 
specified a prior family. However, the procedure corresponds to empirical Bayes if we identify 
$q(\bftheta | \bfpsi^*) = p(\bftheta | \bfx , \bfpsi)$. Thus we call our method AVO-EB and note that it is applicable even outside of the likelihood-free context.

Algorithm~\ref{alg:avo}, can be easily modified to have different mini-batch $M_r$ for the observed data and $M_g$ for the simulator. 
When $M_g \gg M_r$, then the variance of $\bftheta^*$ scales ad $\sqrt{M_r}$~\citep{bernton2017inference}. A final $\sqrt{N/M}$ scaling to the log-posterior $q(\bftheta | \bfpsi)$ is needed to reflect the full strength of the data.
